Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-05-05 paper.qmd
Fixed some typos and made one sentence passive

With forecasts predicting a steadily growing number of flights, see
@fig-forecast, and the modern race to commercialise space, there is a
need to re-evaluate the risks of collision and casualties due to
uncontrolled debris re-entries. The increasing risk landscape poses
challenges in terms of oversight and risk management. The operational and 
economic demand by air transport requires a tradeoff with large-scale
airspace and flow management measures to address the inherent safety risk.
There is a growing demand by overseeing authorities for new methodologies and tools to
assess which safety measures are the most adequate to minimise potential
casualties and economic impact.

The increase in space operations does not exist in a vacuum. The
continual growth of air traffic - despite temporal bumps - is evidenced
by the surging number of flights in Europe [@forecastUpdate2025] and
across the world. On-going development/deployment activities of
mega-constellations of satellites in \acr{LEO}, such as the
already active Starlink and OneWeb or planned ones (e.g. Kuiper,
Starshield, Xingwang), will add thousand of orbiting objects with a short
operational life. Some of these will inevitably re-enter the Earth's
atmosphere in an uncontrolled manner, c.f. [@tbl-megaconstellations;
@megaconstellations].
These numbers and occurence rates will pose challenges to the steadily increasing air traffic.

@@

The modelling and tracking of uncontrolled re-entries poses challenges in terms of safety awareness and operational risk.
Actual and historic data is sparse, and the size, shape, and attitude or
rotation of re-entering objects increase the level of uncertainty for
modelling. To support the decision-making, common approaches employ a mix of handling the uncertainty of
object or debris movement and the associated exposed area for encounter, including
temporal uncertainty windows.
Thus, computationally efficient methods to assess and continously revalidate the collision risk of uncontrolled re-entries and aircraft will help to fine-tune the decision-making and geo-spactial and temporal constraints on aviation.
```         
  <!-- is there more to say and provide context? -->
```

@@

If we consider historical uncontrolled debris re-entry data, we can numerically
derive the debris impact probability as a weighting function by summing up and
averaging all single densities:

>>>>

With forecasts predicting a steadily growing number of flights, see
@fig-forecast, and the modern race to commercialise space, there is a
need to re-evaluate the risks of collision and casualties due to
uncontrolled debris re-entries. The increasing risk landscape poses
challenges in terms of oversight and risk management. The operational and 
economic demand by air transport requires a trade-off with large-scale
airspace and flow management measures to address the inherent safety risk.
There is a growing demand by overseeing authorities for new methodologies and tools to
assess which safety measures are the most adequate to minimise potential
casualties and economic impact.

The increase in space operations does not exist in a vacuum. The
continual growth of air traffic - despite temporal bumps - is evidenced
by the surging number of flights in Europe [@forecastUpdate2025] and
across the world. On-going development/deployment activities of
mega-constellations of satellites in \acr{LEO}, such as the
already active Starlink and OneWeb or planned ones (e.g. Kuiper,
Starshield, Xingwang), will add thousand of orbiting objects with a short
operational life. Some of these will inevitably re-enter the Earth's
atmosphere in an uncontrolled manner, c.f. [@tbl-megaconstellations;
@megaconstellations].
These numbers and occurrence rates will pose challenges to the steadily increasing air traffic.

@@

The modelling and tracking of uncontrolled re-entries poses challenges in terms of safety awareness and operational risk.
Actual and historic data is sparse, and the size, shape, and attitude or
rotation of re-entering objects increase the level of uncertainty for
modelling. To support the decision-making, common approaches employ a mix of handling the uncertainty of
object or debris movement and the associated exposed area for encounter, including
temporal uncertainty windows.
Thus, computationally efficient methods to assess and continuously revalidate the collision risk of uncontrolled re-entries and aircraft will help to fine-tune the decision-making and geospatial and temporal constraints on aviation.

<!-- is there more to say and provide context? -->

@@

Considering historical uncontrolled debris re-entry data, we can numerically
derive the debris impact probability as a weighting function by summing up and
averaging all single densities:
